\(Q\)-Bosons and their relations with integrable hierarchies

### 2024-04-19 draft.tex
finish bulk of the draft

@@

Therefore, the states \(\ket{\vec{n}} \in \mathcal{F}_Q\) are defined as 
\begin{equation}
  \ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots \otimes \ket{n_M} 
 : =  (b_0^\dagger)^{n_0} (b_1^\dagger)^{n_1} \cdots  (b_M^\dagger)^{n_M} \ket{\bm{0}} \; ,
\end{equation}
where we write \(\phi_i^\dagger \equiv \bm{1} \otimes  \cdots \otimes
\phi_i^\dagger \otimes \cdots \otimes \bm{1}\).
Moreover, the norm of these states satisfy 
\begin{equation}
  \bracket{\vec{n}}{\vec{m}} = \frac{[m_0]!}{\prod_{i=1}^M[m_i]!} \delta_{\vec{n}\vec{m}}\qquad 
    [n]! =
    \left\{
    \begin{array}{ll}
    \prod_{i=1}^n(1 - Q^i) & \textrm{if}\ \ n\neq 0 \\
    1 & \textrm{otherwise}
  \end{array}\right.\; .
\end{equation}
As before, we can associate to a given states \(\ket{\vec{n}} =
\ket{n_0, n_1, \dots, n_M}\), a partition -- or Young diagram --
\(\lambda = (1^{n_1} 2^{n_2} \cdots M^{n_M})\), but now we assume that 
there is a proportionaly  factor relating these two objects, 
\begin{equation}
  \ket{\lambda}  = b_\lambda(Q) \ket{\vec{n}}  \ \; , \qquad b_\lambda(Q) = \prod_i [p_i(\lambda)]! 
\end{equation}
where \(p_i(\lambda)\) denotes the number of parts of size \(i\) in
the partition. These partition states satisfy \(
\bracket{\lambda}{\mu} = b_\lambda(Q) \delta_{\lambda,\mu}\). Once
again, we see that the correspondence is not unique since the number
of oscilators at the site \(i=\) are completely ignored in the
partition notation.  An finally, if the number of particles is fixed,
then \(n_0 = N - \ell(\lambda)\).

@@

Bogoliubov~\cite{Bogoliubov2005}, has shown that the scalar product of
two vectors in the \(N\)-particles sector in a chain of length \(M+1\)
is
\begin{equation}
\begin{split}
\label{eq:scalar}
  S(N, M|\bm{x}, \bm{y}) & =
  \bra{\bm{0}} \prod_{j=1}^N \mathbb{C}(x_j) \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}} \\ 
 & = \frac{ \det H(\bm{x},\bm{y})}{ \prod_{i<j}(x_i - x_j)(y_i - y_j)} \; ,
\end{split}
\end{equation}
where \(H\) is an \(N\times N\) matrix with components
\begin{equation}
\label{eq:h-matrix}
  H_{ij} = H(x_i, y_j) 
  =\frac{1 - (x_i y_j)^{ M + N}}{1 - x_i y_j }\; .
\end{equation}

@@

\subsection{Toda tau functions}
In this section, we argue that the scalar product defined above is a tau function 
of the Toda hierarchy. Let us first write the function \(H(z,w)\) as the geometric sum 
\begin{equation}
  H(z,w) = \frac{1 - (zw)^{M+N}}{1 - zw} = \sum_{k=1}^{M+N} (zw)^{k-1} \; , 
\end{equation}
in such a way that 
we can write
\begin{equation}
  \det_{i,j} \left(H(x_i, y_j)\right) = \det_{i,j} \left( \sum_{k=1}^{M+N} x_i^{k-1} y_j^{k-1}\right)
\end{equation}
and we can thing of this expression as the product of a \(N\times (N+M)\) matrix \(\mathcal{X}\)
and another \((M + N)\times N\) matrix \(\mathcal{Y}\), given by
\begin{equation}
  \mathcal{X} = 
  \begin{pmatrix}
  x_1^0 & x_1^1 & \dots & x_1^{M+N-1} \\  
  x_2^0 & x_2^1 & \dots & x_2^{M+N-1} \\  
  \vdots \\
  x_N^0 & x_N^1 & \dots & x_N^{M+N-1} 
  \end{pmatrix}\quad \textrm{and} \quad 
  \mathcal{Y} = 
  \begin{pmatrix}
  y_1^0 & y_2^0 & \dots & y_N^0 \\  
  y_1^1 & y_2^1 & \dots & y_N^1 \\  
  \vdots & \vdots & & \vdots \\
  y_1^{N+M-1} & y_2^{M+N-1} & \dots & y_N^{M+N-1} 
  \end{pmatrix}
\end{equation}
therefore 
\begin{equation}
  \det_{i,j} \left(H(x_i, y_j)\right) =  \det_{i,j} \left( \mathcal{X}\mathcal{Y}\right)
= \sum_{0 \leq \ell_{N} \leq \dots \leq \ell_1\leq N+M }  \det_{ik}(x_i^{\ell_k})  \det_{ik}(y_j^{\ell_k})\; .
\end{equation}
If we now define \(\ell_j = \lambda_k - k + N\), and
using~(\ref{eq:scalar}), we have
\begin{equation}
\label{eq:scalar_exp}
S(N, M|\bm{x}, \bm{y}) = \sum_{\lambda\subseteq [N,M]}
\frac{\det_{ik}(x_i^{\lambda_k - k + N})}{\Delta(\bm{x})} \frac{ \det_{ik}(y_j^{\lambda_k - k + N})}{\Delta(\bm{y})}
= \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{x}) s_\lambda(\bm{y}) \; .
\end{equation}
where \(\Delta(\bm{x})\) and \(\Delta(\bm{x})\) are Vandermonde
determinants. This formula agrees with the Schur expansion defined
in~\cite{Bogoliubov2005}.

Let us define two sets of Miwa coordinates \(\bm{t} = (t_1, t_2, \dots)\)
and \(\bm{t}' = (t'_{-1}, t'_{-2}, \dots)\) by
\begin{equation}
  t_p = \frac{1}{p}\sum_{j=1}^N x_j^p\qquad 
  t'_{-p} = \frac{1}{p}\sum_{j=1}^N y_j^p\; ,
\end{equation}
therefore, we have 
\begin{equation}
  S(N, M|\bm{t}, \bm{t}') = \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
\end{equation}
This expression is known to be a tau function for \(M, N\to \infty\).
In fact, using the free fermions representation~\cite{Alexandrov:2012tr},
it can be written as
\begin{equation}
\sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
 = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} e^{-\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
\end{equation}
It is a tau function of the Toda hierarchy with trivial element \(1 \in GL(\infty)\).
In fact, it is nothing but the Cauchy's identity
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
    = \exp \left( \sum_{m\geq1} m t_m t'_{-m} \right) \; .
\end{equation}
where \(\bm{t}\) and \(\bm{t}'\) are two independent sets of Miwa coordinates.

@@

As a final remark, we can also write
\begin{equation}
  H(z,w) = \sum_{k=0}^{M+N} h_k(zw)^{k-1} \; , 
\end{equation}
with \(h_k=1\) if \(k\in [0, M+N-1]\) and \(0\) otherwise. In this
case, one can define a diagonal \((N+M)\times (N+M)\) matrix
\(\mathcal{H} = \textrm{diag}(h_0, \dots, h_{M+N-1})\). Therefore, if
we repeat the arguments above, we find
\begin{equation}
S(N, M|\bm{t}, \bm{t}') = \sum_{\lambda} h_\lambda s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
\end{equation}
where \(h_\lambda = 1\) if \(\lambda \subset [N,M]\) and it vanishes
otherwise.  In this case, we have that this tau function is of the
type considered in~\cite{orlov:2001}. Evidently that it is a trivial
example, but we expect that in the case of the \(q\)-Boson model,
where \(q\) is finite, the diagonal terms \(h_\lambda\) are more
interesting. We will discuss this case next section.

@@

It is also interesting to notice that the particular case when \(M\to
\infty\), but with a finite number of particles \(N\). The
partitions \(\lambda \in [N, \infty]\) satisfy the condition
\(\ell(\lambda) \leq N\). Therefore, the scalar
product~(\ref{eq:scalar_exp}) becomes
\begin{equation}
  S_N(\bm{t}, \bm{t}') = \sum_{\substack{\lambda \\ \ell(\lambda) \leq N}}
  s_\lambda(\bm{t})  s_\lambda(\bm{t}') \; .
\end{equation}
From~\cite{Zabrodin:2010ii}, we know that this expression can be
written as the following integral
\begin{equation}
  S_N(\bm{t}, \bm{t}') =
  \frac{1}{N!} \prod_{\ell=1}^N \oint_{\Gamma_\ell} \frac{dz_\ell}{2 \pi i z_\ell}
  e^{\xi(\bm{t}, z_\ell) - \xi(\bm{t}', z_{\ell}^{-1})} \Delta(z)\Delta(z^{-1})\; ,
\end{equation}
where
\begin{equation}
\xi(\bm{t}, z) = \sum_{k\geq 0} t_k z^k \qquad
\xi(\bm{t}', 1/z) = \sum_{k\geq 0} t'_{-k} z^{-k} \; . 
\end{equation}
See also~\cite{Kharchev:1991gd} for a detailed proof of this relation,
and~\cite{Orlov:2005} for other details.

@@

\subsubsection{KP tau function}
Finally, one may also observe that if we fix one set of coordinates,
say \(\bm{y}\), then we write
\begin{equation}
 S(N,M|\bm{t}) 
 = \sum_{\lambda} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
 \equiv \sum_{\lambda} c_\lambda(\bm{y})  s_\lambda(\bm{t}) \; ,
\end{equation}
with coefficients \(c_\lambda(\bm{y}) = \det (h_{\lambda_i-i +j}(\bm{y}))\).
But now, it is a trivial to notice that these are Plücker
coordinates in the Jacobi-Trudi form, see~\cite{Miwa2000, Alexandrov:2012tr}.

Therefore, the expression
\begin{equation}
 S(N,M|\bm{t}) = \sum_{\lambda \subseteq [N,M]} s_\lambda(\bm{y})  s_\lambda(\bm{t}) \; ,
\end{equation}
is also a KP tau function: A fact that we already know
from~\cite{Wheeler:2010vmq}, where the author has proved this
statement using the free fermions formalism.

@@

\subsection{Correlation functions}
In~\cite{Bogoliubov2005}, Bogoliubov has also shown that the
correlation functions
\begin{subequations}
\begin{equation}
\begin{split}
\label{eq:correlation}
  P_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
  & = \bra{0} \prod_{j=1}^N \mathbb{C}(x_j)
  \prod_{k=1}^{N-1} \mathbb{B}(y_k) \phi_m^\dagger \ket{0}\\
  & =  \prod_{j=1}^N x_{j}^{M/2} \prod_{k=1}^{N-1} y_{j}^{M/2}
  \bra{0} \prod_{j=1}^N C(1/x_j) \prod_{k=1}^{N-1} B(y_k) \phi_m^\dagger \ket{0}
\end{split}
\end{equation}
can be written as
\begin{equation}
  P_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\}) = 
  \frac{(-1)^{N-1}}{y_N^{(N-1)/2}} \prod_{j=1}^N x_{j}^{M/2}
  \prod_{k=1}^{N-1} y_{j}^{M/2}
  \left( \prod_{t<N} \frac{y_N - y_t}{y_t} \right) \frac{\det Q}{\det H} S(N,M|\bm{x}, \bm{y})
\end{equation}
\end{subequations}
where the \(N\times N\) matrix \(Q\) has components 
\begin{equation}
 Q_{jN} = x_j^{(M + N - 1- 2m)/2} \quad  \textrm{and} \quad 
 Q_{jk} = H_{jk} \; , 
\end{equation}
where \(H_{jk} = H(x_j, y_k)\) are the components of the matrix \(H\) in~(\ref{eq:h-matrix}). 
Since the components \(Q_{jN}\) are independent of the coordinates \(y\), 
we cannot write the above expression as a Toda hierarchy tau
function.

We already know from~(\ref{eq:scalar}) that 
\begin{equation}
   S(N,M|\bm{x}, \bm{y}) =  \Delta(x) \Delta(y) \det H \; ,
\end{equation}
therefore we have
\begin{equation}
  P_m(N, M|\bm{x}, \bm{y}\setminus{y_N})  \propto
  \frac{\det Q}{\Delta(x)} \equiv \mathcal{P}_m(N,M|\bm{x}, \bm{y})\; ,
\end{equation}
and we are going to treat the coordinates \(\{ \bm{y} \}\) as
a set of \(N\) fixed parameters. Furthermore, we write \(N\)
functions \(\bm{F}(z) = (F_1, \dots, F_N)\) defined as
\begin{equation}
  F_j (z) = H(z, y_j) \quad \textrm{if} \ j \neq N\; , \qquad \textrm{and}\qquad 
   F_N (z)  = z^{(M + N - 1 - 2m)/2} \; .
\end{equation}
As before, we expand \(F_j(z)\) as the geometric sum
\begin{equation}
  F_j(z) = \frac{1 - (y_j z)^{M + N}}{1 - y_j z} = \sum_{n=0}^{N + M - 1} (y_j z)^n
  \equiv \sum_{n=0}^{M + N -1} f_{j, n} z^n\; , 
\end{equation}
where we will also assume the \(M + N - 1 \in 2 \mathbb{Z}\).

With these definitions, we conclude that
\begin{equation}
\mathcal{P}_m(N,M|\bm{x}, \bm{y}) =\frac{\det_{jk} F_j(x_k)}{\Delta(x)}\; .
\end{equation}
From the expansion 
\begin{equation}
  \det_{jk} F_j(x_k) = \det_{jk} \left(  \sum_{n=0}^{M + N -1} f_{j, n} x_k^n \right) \; .
\end{equation}
and with the aid of the Cauchy-Binet formula, we write  
\begin{equation}
    \mathcal{P}_m(N,M|\bm{x}, \bm{y})
  = \sum_{0\leq \ell_N\leq \dots \leq \ell_1 \leq N+M}
  \frac{\det_{jk}(f_{j, \ell_k}) \det_{jk}(x_k^{\ell_j})}{\Delta(x)}\; . 
\end{equation}
Now we basically use the definition of the Schur polynomials and the
Jacobi-Trudi expression of Plücker coordinates, see~\cite{Alexandrov:2012tr}. 
Finally, this last expression becomes
\begin{equation}
\mathcal{P}_m(N,M|\bm{x}, \bm{y}) =
\sum_{\lambda} c_\lambda(\bm{y}) s_\lambda(\bm{x}) \; ,
\end{equation}
where \(c_\lambda \equiv \det_{jk}(y_j^{\ell_k})\),  \(\ell_k = \lambda_k - k +N\). Therefore,
we have that this expression is a KP tau function.

From the coordinate expansion of the Bethe
vectors~\cite{Bogoliubov2005, Tsilevich:2006}, it is easy to show that
\begin{equation}
\prod_{j=1}^N \mathbb{B}(x_j)\ket{\lambda}  = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\ket{\mu}\qquad 
 \bra{\lambda} \prod_{j=1}^N \mathbb{C}(x_j) = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\bra{\mu}
\end{equation}
where \(s_{\mu/\lambda}\) are Schur polynomials~\cite{Macdonald:1998}.
Therefore, we can write the correlation function~(\ref{eq:correlation}) as 
\begin{equation}
\begin{split}
  P_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
  & = \sum_\lambda \sum_{\mu \supset (m)} s_{\lambda}(\bm{x}) s_{\mu/(m)}(\bm{y}\setminus\{y_N\})\\
  & = \sum_\mu s_{\mu/(m)}(\bm{y}\setminus\{y_N\}) s_{\mu}(\bm{x})\; .
\end{split}
\end{equation}
where we have used that \(\phi^\dagger \ket{\bm{0}} = \ket{(m)}\), and in the second line
we sum over all partitions, since \(s_{\mu/(m)} = 0\), \(\forall \) \(\mu \not \supset (m)\).
This expression also shows that the expression \(P_m\) can also be written as a tau function.

>>>>

%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
\section{Introduction}
%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*

@@

Therefore, the states \(\ket{\vec{n}} \in \mathcal{F}_Q\) are defined as 
\begin{equation}
  \ket{\vec{n}} = \ket{n_0}\otimes \ket{n_1} \otimes \cdots \otimes \ket{n_M} 
 : =  (b_0^\dagger)^{n_0} (b_1^\dagger)^{n_1} \cdots  (b_M^\dagger)^{n_M} \ket{\bm{0}} \; ,
\end{equation}
where we write \(\phi_i^\dagger \equiv \bm{1} \otimes  \cdots \otimes
\phi_i^\dagger \otimes \cdots \otimes \bm{1}\).
Moreover, the norm of these states satisfy 
\begin{equation}
  \bracket{\vec{n}}{\vec{m}} = \frac{[m_0]!}{\prod_{i=1}^M[m_i]!} \delta_{\vec{n}\vec{m}}\qquad 
    [n]! =
    \left\{
    \begin{array}{ll}
    \prod_{i=1}^n(1 - Q^i) & \textrm{if}\ \ n\neq 0 \\
    1 & \textrm{otherwise}
  \end{array}\right.\; .
\end{equation}
As before, we can associate to a given states \(\ket{\vec{n}} =
\ket{n_0, n_1, \dots, n_M}\), a partition -- or Young diagram --
\(\lambda = (1^{n_1} 2^{n_2} \cdots M^{n_M})\), but now we assume that 
there is a proportionaly  factor relating these two objects, 
\begin{equation}
  \ket{\lambda}_Q  = b_\lambda(Q) \ket{\vec{n}}  \ \; , \qquad b_\lambda(Q) = \prod_i [p_i(\lambda)]! 
\end{equation}
where \(p_i(\lambda)\) denotes the number of parts of size \(i\) in
the partition. These partition states satisfy \(
\bracket{\lambda}{\mu}_Q = b_\lambda(Q) \delta_{\lambda,\mu}\). Once
again, we see that the correspondence is not unique since the number
of oscilators at the site \(i=0\) are completely ignored in the
partition notation. And finally, if the number of particles is fixed,
then \(n_0 = N - \ell(\lambda)\).

@@

Bogoliubov~\cite{Bogoliubov2005}, has shown that the scalar product of
two vectors in the \(N\)-particles sector in a chain of length \(M+1\)
is
\begin{equation}
\begin{split}
\label{eq:scalar}
  \mathcal{I}(N, M|\bm{x}, \bm{y}) & =
  \bra{\bm{0}} \prod_{j=1}^N \mathbb{C}(x_j) \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}} \\ 
 & = \frac{ \det H(\bm{x},\bm{y})}{ \prod_{i<j}(x_i - x_j)(y_i - y_j)} \; ,
\end{split}
\end{equation}
where \(H\) is an \(N\times N\) matrix with components
\begin{equation}
\label{eq:h-matrix}
  H_{ij} = H(x_i, y_j) 
  =\frac{1 - (x_i y_j)^{ M + N}}{1 - x_i y_j }\; .
\end{equation}

@@

\subsection{Toda tau functions}
In this section, we argue that the scalar product defined above is a tau function 
of the Toda hierarchy. Let us first write the function \(H(z,w)\) as the geometric sum 
\begin{equation}
\label{eq:h-exp}
  H(z,w) = \frac{1 - (zw)^{M+N}}{1 - zw} = \sum_{k=1}^{M+N} (zw)^{k-1} \; , 
\end{equation}
in such a way that 
we can write
\begin{equation}
  \det_{i,j} \left(H(x_i, y_j)\right) = \det_{i,j} \left( \sum_{k=1}^{M+N} x_i^{k-1} y_j^{k-1}\right)
\end{equation}
and we can thing of this expression as the product of a \(N\times (N+M)\) matrix \(\mathcal{X}\)
and another \((M + N)\times N\) matrix \(\mathcal{Y}\), given by
\begin{equation}
  \mathcal{X} = 
  \begin{pmatrix}
  x_1^0 & x_1^1 & \dots & x_1^{M+N-1} \\  
  x_2^0 & x_2^1 & \dots & x_2^{M+N-1} \\  
  \vdots \\
  x_N^0 & x_N^1 & \dots & x_N^{M+N-1} 
  \end{pmatrix}\quad \textrm{and} \quad 
  \mathcal{Y} = 
  \begin{pmatrix}
  y_1^0 & y_2^0 & \dots & y_N^0 \\  
  y_1^1 & y_2^1 & \dots & y_N^1 \\  
  \vdots & \vdots & & \vdots \\
  y_1^{N+M-1} & y_2^{M+N-1} & \dots & y_N^{M+N-1} 
  \end{pmatrix}
\end{equation}
therefore 
\begin{equation}
  \det_{i,j} \left(H(x_i, y_j)\right) =  \det_{i,j} \left( \mathcal{X}\mathcal{Y}\right)
= \sum_{0 \leq \ell_{N} \leq \dots \leq \ell_1\leq N+M }  \det_{ik}(x_i^{\ell_k})  \det_{ik}(y_j^{\ell_k})\; .
\end{equation}
If we now define \(\ell_j = \lambda_k - k + N\), and
using~(\ref{eq:scalar}), we have
\begin{equation}
\label{eq:scalar_exp}
\mathcal{I}(N, M|\bm{x}, \bm{y}) = \sum_{\lambda\subseteq [N,M]}
\frac{\det_{ik}(x_i^{\lambda_k - k + N})}{\Delta(\bm{x})} \frac{ \det_{ik}(y_j^{\lambda_k - k + N})}{\Delta(\bm{y})}
= \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{x}) s_\lambda(\bm{y}) \; .
\end{equation}
where \(\Delta(\bm{x})\) and \(\Delta(\bm{x})\) are Vandermonde
determinants. This formula agrees with the Schur expansion defined
in~\cite{Bogoliubov2005}.

Let us define two sets of Miwa coordinates \(\bm{t} = (t_1, t_2, \dots)\)
and \(\bm{t}' = (t'_{-1}, t'_{-2}, \dots)\) by
\begin{equation}
  t_p = \frac{1}{p}\sum_{j=1}^N x_j^p\qquad 
  t'_{-p} = \frac{1}{p}\sum_{j=1}^N y_j^p\; ,
\end{equation}
therefore, we have 
\begin{equation}
  \mathcal{I}(N, M|\bm{t}, \bm{t}') = \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
\end{equation}
This expression is known to be a tau function for \(M, N\to \infty\).
In fact, using the free fermions representation~\cite{Alexandrov:2012tr},
it can be written as
\begin{equation}
\sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
 = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} e^{-\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
\end{equation}
It is a tau function of the Toda hierarchy with trivial element \(1 \in GL(\infty)\).
In fact, it is nothing but the Cauchy's identity
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
    = \exp \left( \sum_{m\geq1} m t_m t'_{-m} \right) \; .
\end{equation}
where \(\bm{t}\) and \(\bm{t}'\) are two independent sets of Miwa coordinates.

@@

As a final remark, we can also write
\begin{equation}
  H(z,w) = \sum_{k=0}^{M+N} h_k(zw)^{k-1} \; , 
\end{equation}
with \(h_k=1\) if \(k\in [0, M+N-1]\) and \(0\) otherwise. In this
case, one can define a diagonal \((N+M)\times (N+M)\) matrix
\(\mathcal{H} = \textrm{diag}(h_0, \dots, h_{M+N-1})\). Therefore, if
we repeat the arguments above, we find
\begin{equation}
\mathcal{I}(N, M|\bm{t}, \bm{t}') = \sum_{\lambda} h_\lambda s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
\end{equation}
where \(h_\lambda = 1\) if \(\lambda \subset [N,M]\) and it vanishes
otherwise.  In this case, we have that this tau function is of the
type considered in~\cite{orlov:2001}. Evidently that it is a trivial
example, but we expect that in the case of the \(q\)-Boson model,
where \(q\) is finite, the diagonal terms \(h_\lambda\) are more
interesting. We will discuss this case next section.

@@

It is also interesting to notice that the particular case when \(M\to
\infty\), but with a finite number of particles \(N\). The
partitions \(\lambda \in [N, \infty]\) satisfy the condition
\(\ell(\lambda) \leq N\). Therefore, the scalar
product~(\ref{eq:scalar_exp}) becomes
\begin{equation}
  \mathcal{I}_N(\bm{t}, \bm{t}') = \sum_{\substack{\lambda \\ \ell(\lambda) \leq N}}
  s_\lambda(\bm{t})  s_\lambda(\bm{t}') \; .
\end{equation}
From~\cite{Zabrodin:2010ii}, we know that this expression can be
written as the following integral
\begin{equation}
  \mathcal{I}_N(\bm{t}, \bm{t}') =
  \frac{1}{N!} \prod_{\ell=1}^N \oint_{\Gamma_\ell} \frac{dz_\ell}{2 \pi i z_\ell}
  e^{\xi(\bm{t}, z_\ell) - \xi(\bm{t}', z_{\ell}^{-1})} \Delta(z)\Delta(z^{-1})\; ,
\end{equation}
where
\begin{equation}
\xi(\bm{t}, z) = \sum_{k\geq 0} t_k z^k \qquad
\xi(\bm{t}', 1/z) = \sum_{k\geq 0} t'_{-k} z^{-k} \; . 
\end{equation}
See also~\cite{Kharchev:1991gd} for a detailed proof of this relation,
and~\cite{Orlov:2005} for other details.

@@

@@

\subsubsection{KP tau function}
Finally, one may also observe that if we fix one set of coordinates,
say \(\bm{y}\), then we write
\begin{equation}
 \mathcal{I}(N,M|\bm{t}) 
 = \sum_{\lambda} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
 \equiv \sum_{\lambda} c_\lambda(\bm{y})  s_\lambda(\bm{t}) \; ,
\end{equation}
with coefficients \(c_\lambda(\bm{y}) = \det (h_{\lambda_i-i +j}(\bm{y}))\).
But now, it is a trivial to notice that these are Plücker
coordinates in the Jacobi-Trudi form, see~\cite{Miwa2000, Alexandrov:2012tr}.

Therefore, the expression
\begin{equation}
 \mathcal{I}(N,M|\bm{t}) = \sum_{\lambda \subseteq [N,M]} s_\lambda(\bm{y})  s_\lambda(\bm{t}) \; ,
\end{equation}
is also a KP tau function: A fact that we already know
from~\cite{Wheeler:2010vmq}, where the author has proved this
statement using the free fermions formalism.

\subsection{Correlation functions}
In~\cite{Bogoliubov2005}, Bogoliubov has also shown that the
correlation functions
\begin{subequations}
\begin{equation}
\begin{split}
\label{eq:correlation}
  A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
  & = \bra{0} \prod_{j=1}^N \mathbb{C}(x_j)
  \prod_{k=1}^{N-1} \mathbb{B}(y_k) \phi_m^\dagger \ket{0}\\
  & =  \prod_{j=1}^N x_{j}^{M/2} \prod_{k=1}^{N-1} y_{j}^{M/2}
  \bra{0} \prod_{j=1}^N C(1/x_j) \prod_{k=1}^{N-1} B(y_k) \phi_m^\dagger \ket{0}
\end{split}
\end{equation}
can be written as
\begin{equation}
  A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\}) = 
  \frac{(-1)^{N-1}}{y_N^{(N-1)/2}} \prod_{j=1}^N x_{j}^{M/2}
  \prod_{k=1}^{N-1} y_{j}^{M/2}
  \left( \prod_{t<N} \frac{y_N - y_t}{y_t} \right) \frac{\det Q}{\det H} S(N,M|\bm{x}, \bm{y})
\end{equation}
\end{subequations}
where the \(N\times N\) matrix \(Q\) has components 
\begin{equation}
 Q_{jN} = x_j^{(M + N - 1- 2m)/2} \quad  \textrm{and} \quad 
 Q_{jk} = H_{jk} \; , 
\end{equation}
where \(H_{jk} = H(x_j, y_k)\) are the components of the matrix \(H\) in~(\ref{eq:h-matrix}). 
Since the components \(Q_{jN}\) are independent of the coordinates \(y\), 
we cannot write the above expression as a Toda hierarchy tau
function.

We already know from~(\ref{eq:scalar}) that 
\begin{equation}
    \frac{\mathcal{I}(N,M|\bm{x}, \bm{y})}{\det H}=  \frac{1}{\Delta(x)\Delta(y)} \; ,
\end{equation}
therefore we have
\begin{equation}
  A_m(N, M|\bm{x}, \bm{y}\setminus{y_N})  \propto
  \frac{\det Q}{\Delta(x)} \equiv \mathcal{A}_m(N,M|\bm{x}, \bm{y})\; ,
\end{equation}
and we are going to treat the coordinates \(\{ \bm{y} \}\) as
a set of \(N\) fixed parameters. Furthermore, we write \(N\)
functions \(\bm{F}(z) = (F_1, \dots, F_N)\) defined as
\begin{equation}
  F_j (z) = H(z, y_j) \quad \textrm{if} \ j \neq N\; , \qquad \textrm{and}\qquad 
   F_N (z)  = z^{(M + N - 1 - 2m)/2} \; .
\end{equation}
As before, we expand \(F_j(z)\) as the geometric sum
\begin{equation}
  F_j(z) = \frac{1 - (y_j z)^{M + N}}{1 - y_j z} = \sum_{n=0}^{N + M - 1} (y_j z)^n
  \equiv \sum_{n=0}^{M + N -1} f_{j, n} z^n\; , 
\end{equation}
where we will also assume the \(M + N - 1 \in 2 \mathbb{Z}\).

With these definitions, we conclude that
\begin{equation}
\mathcal{A}_m(N,M|\bm{x}, \bm{y}) =\frac{\det_{jk} F_j(x_k)}{\Delta(x)}\; .
\end{equation}
From the expansion 
\begin{equation}
  \det_{jk} F_j(x_k) = \det_{jk} \left(  \sum_{n=0}^{M + N -1} f_{j, n} x_k^n \right) \; .
\end{equation}
and with the aid of the Cauchy-Binet formula, we write  
\begin{equation}
    \mathcal{A}_m(N,M|\bm{x}, \bm{y})
  = \sum_{0\leq \ell_N\leq \dots \leq \ell_1 \leq N+M}
  \frac{\det_{jk}(f_{j, \ell_k}) \det_{jk}(x_k^{\ell_j})}{\Delta(x)}\; . 
\end{equation}
Now we basically use the definition of the Schur polynomials and the
Jacobi-Trudi expression of Plücker coordinates, see~\cite{Alexandrov:2012tr}. 
Finally, this last expression becomes
\begin{equation}
\mathcal{A}_m(N,M|\bm{x}, \bm{y}) =
\sum_{\lambda} c_\lambda(\bm{y}) s_\lambda(\bm{x}) \; ,
\end{equation}
where \(c_\lambda \equiv \det_{jk}(y_j^{\ell_k})\),  \(\ell_k = \lambda_k - k +N\). Therefore,
we have that this expression is a KP tau function.

From the coordinate expansion of the Bethe
vectors~\cite{Bogoliubov2005, Tsilevich:2006}, it is easy to show that
\begin{equation}
\prod_{j=1}^N \mathbb{B}(x_j)\ket{\lambda}  = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\ket{\mu}\qquad 
 \bra{\lambda} \prod_{j=1}^N \mathbb{C}(x_j) = \sum_{\mu \supset \lambda } s_{\mu/\lambda}(\bm{x})\bra{\mu}
\end{equation}
where \(s_{\mu/\lambda}\) are Schur polynomials~\cite{Macdonald:1998}.
Therefore, we can write the correlation function~(\ref{eq:correlation}) as 
\begin{equation}
\begin{split}
  A_m(N, M|\bm{x}, \bm{y}\setminus \{y_N\})
  & = \sum_\lambda \sum_{\mu \supset (m)} s_{\lambda}(\bm{x}) s_{\mu/(m)}(\bm{y}\setminus\{y_N\})\\
  & = \sum_\mu s_{\mu/(m)}(\bm{y}\setminus\{y_N\}) s_{\mu}(\bm{x})\; .
\end{split}
\end{equation}
where we have used that \(\phi^\dagger \ket{\bm{0}} = \ket{(m)}\), and in the second line
we sum over all partitions, since \(s_{\mu/(m)} = 0\), \(\forall \) \(\mu \not \supset (m)\).
This expression also shows that the expression \(P_m\) can also be written as a tau function.

\subsubsection{Skew Schur polynomials expansion}
More generally, let us consider the correlation functions 
\begin{equation}
\begin{split}
  A_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y}) & =
  \bra{\lambda_1} \prod_{j=1}^N \mathbb{C}(x_j)
  \prod_{k=1}^{N'} \mathbb{B}(y_k)\ket{\lambda_2} \\
  & = \sum_{\mu} s_{\mu/\lambda_1}(\bm{x}) s_{\mu/\lambda_2}(\bm{y}) \; .
\end{split}
\end{equation}
where \(N' + \ell(\lambda_2) + n_0 = N' + \ell(\lambda_2) + n'_0\) and 
we have also used that the expression for any Young diagram
\(\mu\) that does not contain \(\lambda_1\) and \(\lambda_2\)
vanishes. Using the elementary property of skew Schur
polynomials~\cite{Macdonald:1998}
\begin{equation}
  \sum_{\mu} s_{\mu/\lambda_1}(\bm{x}) s_{\mu/\lambda_2}(\bm{y}) = \prod_{i,j}\frac{1}{1 - x_i y_j}
 \sum_\nu s_{\lambda_1/\nu}(\bm{x}) s_{\lambda_2/\nu}(\bm{y})\; ,
\end{equation}
and the Cauchy's identity, we have 
\begin{equation}
\begin{split}
  A_{\lambda_1 \lambda_2}(N', N, M|\bm{x}, \bm{y}) & =
 \sum_\mu s_{\mu}(\bm{x}) s_{\mu}(\bm{y})
 \sum_\nu s_{\lambda_1/\nu}(\bm{x}) s_{\lambda_2/\nu}(\bm{y})\; .
\end{split}
\end{equation}
Therefore, this correlation function is nothing but the off-shell
norm~(\ref{eq:scalar}) times a finite sum over skew Schur functions.
This expression shows that the existence of these tau functions is not
a trivial fact of this model.
